Likelihoods for a general class of ARGs under the SMC

### 2025-04-23 paper.tex
Update

% coalescent introduction: backwards-in-time
% left-to-right: Wiuf-Hein (difficult)
% SMC: although easy to reason about left-to-right sense
%
The coalescent process describes the ancestry of a set of sampled genomes.
The original algorithm
to simulate the coalescent with
recombination was formulated backwards-in-time \revpoint{2}{24}
by Hudson \citeyearpar{hudson_properties_1983}.
Wiuf and Hein later described
a tree-by-tree method to simulate the same stochastic process
\citep{wiuf_recombination_1999}.
This along-the-genome approach is considerably more complex as the next
tree to be simulated depends on all previous trees.
This idea however provided the basis of the
SMC \citep{mcvean_approximating_2005}, restricting the set of all state
space transitions possible under the Hudson coalescent to
obtain a process that is Markovian both backwards-in-time,
as well as along-the-genome. \revpoint{2}{25}
In its backwards-in-time formulation, the SMC only requires a
simple modification to the coalescent with recombination.
Instead of allowing any pair of lineages to coalesce,
the SMC is the process in which only lineages with \emph{overlapping}
genomic intervals may coalesce. The SMC' is a refinement
of the SMC which also allows lineages with abutting genomic intervals \revpoint{2}{26}
to coalesce~\citep{marjoram_fast_2006}.
Although \citet{mcvean_approximating_2005} verbally described the SMC
using this backwards-in-time description, their analysis
was almost entirely in left-to-right terms,
and subsequent papers followed this trend
\citep{li_inference_2011,paul_accurate_2011,schiffels_inferring_2014,
rasmussen_genome-wide_2014}.

>>>>

% coalescent introduction: backwards-in-time
% left-to-right: Wiuf-Hein (difficult)
% SMC: although easy to reason about left-to-right sense
%
The coalescent process describes the ancestry of a set of sampled genomes.
The original algorithm
to simulate the coalescent with
recombination was formulated backwards-in-time \revpoint{2}{24}
by Hudson \citeyearpar{hudson_properties_1983}.
Wiuf and Hein later described
a tree-by-tree method to simulate the same stochastic process
\citep{wiuf_recombination_1999}.
This along-the-genome approach is considerably more complex as the next
tree to be simulated depends on all previous trees.
This idea however provided the basis of the
SMC \citep{mcvean_approximating_2005}, restricting the set of all state
space transitions possible under the Hudson coalescent to
obtain a process that is Markovian both backwards-in-time,
as well as along-the-genome. \revpoint{2}{25}
In its backwards-in-time formulation, the SMC only requires a
simple modification to the coalescent with recombination.
Instead of allowing any pair of lineages to coalesce,
the SMC is the process in which only lineages with \emph{overlapping}
genomic intervals may coalesce. The SMC' is a refinement
of the SMC which also allows lineages with adjoining genomic intervals \revpoint{2}{26}
to coalesce~\citep{marjoram_fast_2006}.
Although \citet{mcvean_approximating_2005} verbally described the SMC
using this backwards-in-time description, their analysis
was almost entirely in left-to-right terms,
and subsequent papers followed this trend
\citep{li_inference_2011,paul_accurate_2011,schiffels_inferring_2014,
rasmussen_genome-wide_2014}.
